---
title: "Ststistcal Inference of Exponentials"
author: "nselvak"
date: "8/24/2020"
output: html_document
---

# Overview 

What will be reported is answers to the questions given and the steps taken to answer the questions given. Mainly, it will be some usual data preprocessing and then the actual simulations and the results  

## Instructions

1. Show the sample mean and compare it to the theoretical mean of the distribution.

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3. Show that the distribution is approximately normal.

## Loading Libraries

```{r Loading Libraries, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
```

## Task

```{r Initial Simulations}
# set seed for reproducability
set.seed(84)

# set lambda to 0.2
lambda <- 0.2

# 40 samples
n <- 40

# 1000 simulations
sims <- 1000

# simulate
sim_expo <- replicate(sims, rexp(n, lambda))

# calculate mean of exponentials
mean_expo <- apply(sim_expo, 2, mean)
```

## Question 1
Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r, analytical mean}
# analytical mean
analytical_mean <- mean(mean_expo)
analytical_mean
```

```{r, theoratical mean}
# Theoratical mean
theory_mean <- 1/lambda
theory_mean
```

```{r, visualization}
# visualization

hist(mean_expo, 
     xlab = "mean",
     main = "Exponential Function Simulations")
abline(v = analytical_mean, col = "red")
abline(v = theory_mean, col = "blue")
```

Analytics mean is `r analytical_mean` and theoretical mean `r theory_mean`. Center of distribution for average of 40 exponentials is very close to the theoretical center of the distribution.

## Question 2

Show how variable it is and compare it to the theoretical variance of the distribution.. 

```{r, SD of Distribution for Analytical}
# standard deviation of distribution
sd_dist <- sd(mean_expo)
sd_dist
```

```{r, SD of Distribution for Theoretical}
# standard deviation from analytical expression
sd_theory <- (1/lambda)/sqrt(n)
sd_theory
```

```{r, Variance of Distribution for Analytical}
# variance of distribution
var_dist <- sd_dist^2
var_dist
```

```{r, Variance of Distribution for Theoretical}
# variance from analytical expression
var_theory <- ((1/lambda)*(1/sqrt(n)))^2
var_theory
```

Standard Deviation of the distribution is `r sd_dist` with the theoretical SD calculated to be `r sd_theory`. The Theoretical variance is calculated as ((1 / ??) * (1/???n))<sup>2</sup> = `r var_theory`. The actual variance of the distribution is `r var_dist`.

## Question 3
Show that the distribution is approximately normal.

```{r, Show Normal Distribution}
xfit <- seq(min(mean_expo), max(mean_expo), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))

hist(mean_expo, breaks=n, prob=T, col="cyan",
     xlab = "means", 
     ylab="density", 
     main="Density of means")
lines(xfit, yfit, pch=22, col="black", lty=5)
```

```{r, Compare Distributions}
# compare the distribution of averages of 40 exponentials to a normal distribution
qqnorm(mean_expo)
qqline(mean_expo, col = 2)
```

Quoting the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.